### 2021-04-17 assertions/doc/thesis-bernhardt/chapters/2-use_case_analysis.tex
Revision chpt 3 assertion language

\section{Interesting formulas and special cases}\label{sec:examples}
This section contains a subset of use-cases that were collected in collaboration with Julian Veigl for an intermediate report \cite{bernhardt_veigel_2020}. The selection of examples covers different types and compositions of quantifiers, demonstrates different meanings of their negations, identifies special cases and discusses their implications for the implementation of distributed assertion checking.

@@

\section{Existential quantification and proofs of validity}\label{sec:existential}
So far, this thesis only considered formulas featuring universal quantification, for which the assertion scheme challenged the validators to find counterexamples. However, distributed assertion checking can not only be used to refute assumptions, but also to prove their validity. This is the case for formulas featuring existential quantification. As an example, consider the assumption that two sets $U$ and $V$ have a non-empty intersection, i.e., $U \cap V \neq \emptyset$. This assumption and its negation can be expressed in predicate logic by 
\begin{equation}\label{eq:intersect}
  (\exists i, \exists j) (0 \le i < |u|) \land (0 \le j < |v|) \Rightarrow u[i] = v[j]
\end{equation}
and
\begin{equation}\label{eq:intersect_neg}
  (\forall i, \forall j) (0 \le i < |u|) \land (0 \le j < |v|) \land u[i] \neq v[j]
\end{equation}
If the condition $u[i] \neq v[j]$ is false for random values of $i$ and $j$, it does not produce a counterexample --- it produces a proof for the validity of the assumption, because it identified an element that is present in both sets. In that case, the validator finding such a proof should publish an approval instead of a veto to the network. After an approval has been posted, and all validators have come to a consensus about the validity of the assumption, the transaction may be included in a block (assuming it doesn't fail due to other reasons). For the blockchain protocol, this implies that distributed refutation and validation have to be implemented as separate modi of assertion checking. \\

@@

\section{Nested quantification}
The previous example showed that universal and existential quantification have to be handled differently. However, further special cases caused by nested quantification have to be considered as well. A nesting of the same type of quantifier, e.g. $\forall i \forall j$ can be considered as trivial. They can be handled by generating a random value for each predicate variable and checking for a counterexample or proof respectively. It becomes less trivial though, if the nesting contains both universal and existential quantifiers. This section examines the cases of $\forall i \exists j$ and $\exists i \forall j$ on the basis of a contract that expects a solution for a boolean satisfiability problem (SAT).

\subsection{Boolean satisfiability problem}\label{sec:sat}
Instances of SAT problems consist of a formula $F$ in propositional logic, and determine if there exists a variable assignment $\mathcal{A}$ to the values $true$ or $false$, s.t. the formula is satisfied ($\mathcal{A}(F) = 1$) \cite{Biere2009}. Modern SAT-solvers expect $F$ to be in conjunctive normal form (CNF), i.e., a conjunction of a set of clauses, where clauses are disjunctions of variables or their negations (called literals) \cite{cnf_math_encycl}. While finding an assignment $\mathcal{A}$ that satisfies $F$ is a NP-complete problem \cite{Biere2009}, verifying whether a given assignment satisfies the formula takes $\mathcal{O}(n*m)$, where $n$ is the number of clauses and $m$ the number of variables (in the worst case, every variable is present as a literal in each clause). In the case of k-SAT problems, where each clause contains at most k literals \cite{Biere2009}, the complexity is reduced to $\mathcal{O}(n)$.

@@

\subsection{Satisfiability in predicate logic - CNF}
Consider a contract that expects a SAT problem in CNF and a proposed assignment, both encoded as described above. Assuming that $F$ is a k-SAT instance, the assumption that $\mathcal{A}$ satisfies $F$ can be expressed in predicate logic as follows:
\begin{equation}\label{eq:cnf_sat}
\begin{aligned}
(\forall i : int, \exists j : int) &(0 \leq i < |F|) \wedge (0 \leq j < k) \\
&\Rightarrow ((isPositive(F[i][j]) \wedge \mathcal{A}[decode(F[i][j])]) \\
&\vee (\neg isPositive(F[i][j]) \wedge \neg \mathcal{A}[decode(F[i][j])])))
\end{aligned}
\end{equation}

@@

\subsection{Satisfiability in predicate logic - DNF}\label{sec:dnf}
The same propositional formula with the reversed quantification nesting $\exists i \forall j$ describes the assertion for the assumption $\mathcal{A}(F) = 1$, where $F$ is given in disjunctive normal form (DNF), i.e., a disjunction of conjunctions \cite{dnf_math_encycl}:
\begin{equation}\label{eq:dnf_sat}
\begin{aligned}
(\exists i : int, \forall j : int) \: &(0 \leq i < |F|) \wedge (0 \leq j < k) \\
&\Rightarrow ((isPositive(F[i][j]) \wedge \mathcal{A}[decode(F[i][j])]) \\
&\vee (\neg isPositive(F[i][j]) \wedge \neg \mathcal{A}[decode(F[i][j])])))
\end{aligned}
\end{equation}
The translation to code can be approached in the same way than its CNF counterpart. Instead of a counterexample, this use-case demands for a proof.

@@

\section{Restriction of use-cases}\todo{Better title?}
The examples given in this chapter have shown that, depending on the type and order of quantification, not all properties can be checked in the same way and need different modi of assertion checking, as well as a distinct translation to target code. This increases the complexity and requirements for an implementation of the off-chain infrastructure and protocol amendment; therefore, as a first iteration, this thesis focuses on providing a design and toolchain for assertions using universal quantification only.

>>>>

\section{Examples and special cases}\label{sec:examples}
This section contains a subset of use-cases that were collected in collaboration with Julian Veigl for an intermediate report \cite{bernhardt_veigel_2020}. The selection of examples covers different types and compositions of quantifiers, demonstrates different meanings of their negations, identifies special cases and discusses their implications for the implementation of distributed assertion checking.

@@

\subsection{Existential quantification - intersection of sets}\label{sec:existential}
So far, this thesis only considered formulas featuring universal quantification, for which the assertion scheme challenged the validators to find counterexamples. However, distributed assertion checking can not only be used to refute assumptions, but also to prove their validity. This is the case for formulas featuring existential quantification. As an example, consider the assumption that two sets $U$ and $V$ have a non-empty intersection, i.e., $U \cap V \neq \emptyset$. This assumption and its negation can be expressed in predicate logic by 
\begin{equation}\label{eq:intersect}
  (\exists i, \exists j) (0 \le i < |u|) \land (0 \le j < |v|) \Rightarrow u[i] = v[j]
\end{equation}
and
\begin{equation}\label{eq:intersect_neg}
  (\forall i, \forall j) (0 \le i < |u|) \land (0 \le j < |v|) \land u[i] \neq v[j]
\end{equation}
If the condition $u[i] \neq v[j]$ is false for random values of $i$ and $j$, it does not produce a counterexample --- it produces a proof for the validity of the assumption, because it identified an element that is present in both sets. In that case, the validator finding such a proof should publish an approval instead of a veto to the network. After an approval has been posted, and all validators have come to a consensus about the validity of the assumption, the transaction may be included in a block (assuming it doesn't fail due to other reasons). For the blockchain protocol, this implies that distributed refutation and validation have to be implemented as separate modi of assertion checking. \\

@@

\subsection{Nested quantification - SAT problems}
The previous example showed that universal and existential quantification have to be handled differently. However, further special cases caused by nested quantification have to be considered as well. A nesting of the same type of quantifier, e.g. $\forall i \forall j$ can be considered as trivial. They can be handled by generating a random value for each predicate variable and checking for a counterexample or proof respectively. It becomes less trivial though, if the nesting contains both universal and existential quantifiers. This section examines the cases of $\forall i \exists j$ and $\exists i \forall j$ on the basis of a contract that expects a solution for a boolean satisfiability problem (SAT). \\

Instances of SAT problems consist of a formula $F$ in propositional logic, and determine if there exists a variable assignment $\mathcal{A}$ to the values $true$ or $false$, s.t. the formula is satisfied ($\mathcal{A}(F) = 1$) \cite{Biere2009}. Modern SAT-solvers expect $F$ to be in conjunctive normal form (CNF), i.e., a conjunction of a set of clauses, where clauses are disjunctions of variables or their negations (called literals) \cite{cnf_math_encycl}. While finding an assignment $\mathcal{A}$ that satisfies $F$ is a NP-complete problem \cite{Biere2009}, verifying whether a given assignment satisfies the formula takes $\mathcal{O}(n*m)$, where $n$ is the number of clauses and $m$ the number of variables (in the worst case, every variable is present as a literal in each clause). In the case of k-SAT problems, where each clause contains at most k literals \cite{Biere2009}, the complexity is reduced to $\mathcal{O}(n)$.

@@

\subsubsection{Satisfiability in predicate logic - CNF}
Consider a contract that expects a SAT problem in CNF and a proposed assignment, both encoded as described above. Assuming that $F$ is a k-SAT instance, the assumption that $\mathcal{A}$ satisfies $F$ can be expressed in predicate logic as follows:
\begin{equation}\label{eq:cnf_sat}
\begin{aligned}
(\forall i : int, \exists j : int) &(0 \leq i < |F|) \wedge (0 \leq j < k) \\
&\Rightarrow ((isPositive(F[i][j]) \wedge \mathcal{A}[decode(F[i][j])]) \\
&\vee (\neg isPositive(F[i][j]) \wedge \neg \mathcal{A}[decode(F[i][j])])))
\end{aligned}
\end{equation}

@@

\subsubsection{Satisfiability in predicate logic - DNF}\label{sec:dnf}
The same propositional formula with the reversed quantification nesting $\exists i \forall j$ describes the assertion for the assumption $\mathcal{A}(F) = 1$, where $F$ is given in disjunctive normal form (DNF), i.e., a disjunction of conjunctions \cite{dnf_math_encycl}:
\begin{equation}\label{eq:dnf_sat}
\begin{aligned}
(\exists i : int, \forall j : int) \: &(0 \leq i < |F|) \wedge (0 \leq j < k) \\
&\Rightarrow ((isPositive(F[i][j]) \wedge \mathcal{A}[decode(F[i][j])]) \\
&\vee (\neg isPositive(F[i][j]) \wedge \neg \mathcal{A}[decode(F[i][j])])))
\end{aligned}
\end{equation}
The translation to code can be approached in the same way than its CNF counterpart. Instead of a counterexample, this use-case demands for a proof.

@@

\section{Restriction of use-cases}\label{sec:restrict}\todo{Better title? -> conclusion?}
\todo{meh}
The examples given in this chapter have shown that, depending on the type and order of quantification, not all properties can be checked in the same way and need different modi of assertion checking. Additionally,  as well as a distinct translation to target code. This increases the complexity and requirements for an implementation of the off-chain infrastructure and protocol amendment; therefore, as a first iteration, this thesis focuses on providing a design and toolchain for assertions using universal quantification only.

### 2021-04-17 assertions/doc/thesis-bernhardt/chapters/3-offline_design.tex
Revision chpt 3 assertion language

\chapter{Generic Off-Chain Design}\label{chap:offline}
The implementation of distributed assertion checking on any blockchain requires some off-chain considerations and infrastructure. The previous chapter defined a formal set of logical formulas amenable for this approach and gave some in-depth examples. As stated before, the proposed implementation in the following chapters is, for the most part, restricted to formulas of predicate logic using universal quantification.\\
The purpose of the off-chain infrastructure is to provide a syntax for stating assertions that check logical formulas, as well as a toolchain to compile these assertions into code that can be originated on the target blockchain. Depending on the individual solution for the respective blockchain protocol, this can either be the original contract extended directly with the assertion code, or as a separate contract that extends the original contract in some other way. This chapter describes the generic part of the off-chain design, which corresponds to the frontend of the compilation pipeline shown in \figref{fig:pipeline_frontend}. As a start, \secref{sec:syntax} describes the concrete syntax for writing assertions and shows how it expresses some of the previously given examples. The stages of the pipeline frontend, including parsing and transforming the original assertions to assertions that check for counterexamples, are described in \secref{sec:pipeline_front}. Lastly, \secref{sec:accuracy} analyses the reliability of distributed assertion checking as proposed and to what extent it affects the costs.

@@

\section{Assertion syntax}\label{sec:syntax}
The assertion syntax is described in two versions by the grammars in extended Backus-Naur form (EBNF) shown in appendix \ref{apx:grammar}. One version implements a prefix and the other one an infix notation. The pipeline currently supports the prefix notation with mandatory parentheses to avoid the need of handling operator precedences. To avoid context switches (at least for developers on Tezos), the assertion syntax references those of OCaml and thereby also Michelson.\\
\lstinputlisting[caption=Assertion contract for checking if a list is sorted, language=Assertion, label={lst:sorted_assertion}]{listings/sorted.tza}
A file containing assertions (henceforth referred to as ``assertion contract'') contains at least one assertion for some function or entrypoint of the original contract (henceforth referred to as ``parent contract''). An assertion begins with a signature consisting of an optional tag denoted by \%, followed by a parameter pattern. The tag can be used for documentation purposes, but may sometimes be needed to indicate which function of the parent contract the assertion should be assigned to by mirroring the entrypoint name. Parameter patterns are similar to patterns in OCaml --- they are templates that match shapes of data structures and binds identifiers to the components of them. The pattern given in the signature must match the parameter type of the entrypoint it refers to. The body is an (optional) nesting of quantifiers and conditions around exactly one assertion. Conditionals can be used to restrict the quantification domains or for some other constraints. For completeness, the existential quantifier is already included in the grammar, however the current version of the pipeline will reject any assertions containing it. \\
In order to make the frontend generic, the assertion grammar constitutes a union of operations and types for the target languages. This also allows the transformation to be oblivious of the target platform and be identical for all backends. The current version of the grammar recognizes all types present in the Tezos VM and a subset of operations that Michelson provides (\secref{sec:tezos} contains an introduction to Tezos and Michelson and goes more into detail about this). Responsible for rejecting any assertions containing unsupported types or operations are the respective backends.\\

@@

\section{Accuracy of the approach}\label{sec:accuracy}
Since the proposed approach implements a probabilistic test, the accuracy as well as the costs incurred by guaranteeing a specified certainty threshold, need to be examined closely. The goal is to identify a formula that, given the domain of a formula in predicate logic, returns an estimate of a lower bound of samples necessary to find an existing counterexample with a probability $p$. From this analysis, it can be derived how effective a formula can be checked on a blockchain with $m$ validators. Sections \ref{sec:coupon} and \ref{sec:prob_threshold} depict the findings from \cite{bernhardt_veigel_2020} and are based on the assumptions that the random generator picks elements independently from a uniform distribution and that there exists exactly one counterexample for the passed parameter.

>>>>

\chapter{Toolchain Implementation - Frontend}\label{chap:offline}
The purpose of the toolchain is to support the proposed model from specifying assertions for Smart Contracts, transforming them to their negation to generating assertion Smart Contracts that can be invoked by the validators. As determined in \secref{sec:restrict}, the current version of the toolchain only supports the processing of assertions featuring universal quantification. The code is available in a GitHub\footnote{\url{https://github.com/proglang/thesis-distributed-assertion-checking}\todo{commit}} repository of the Chair of Programming Languages.

As an overview, \figref{fig:pipeline_frontend} depicts all stages of the processing pipeline implemented by the toolchain. The colourized stages highlight the frontend, which was designed to be as generic as possible, s.t. it can be used (or extended) to support any target blockchain. For the target-specific processing, the intermediate result of the frontend is passed to the respective backends via a public interface.

First, developers can specify assertions in a domain-specific language, which is described in \secref{sec:syntax}. The toolchain then parses the assertions into abstract syntax trees (ASTs), which are passed to the transformation stage. In this stage, the ASTs are transformed into ASTs that represent their negation. This is described further in \secref{sec:transformation}. The resulting ASTs are then passed to the backend, which is introduced in the subsequent chapter for the Tezos blockchain.
%Lastly, \secref{sec:accuracy} analyses the reliability of distributed assertion checking as proposed and to what extent it affects the costs.

@@

\section{Assertion syntax}\label{sec:syntax}
Tezos' Smart Contract language Michelson is a low-level and stack-based language. Specifying assertions that way is complex and inconvenient, especially if the language is compiled to other targets as well. Therefore, this thesis defines a domain-specific high-level language to express assertions.

The syntax of this language is described in two versions by the EBNF grammars given in appendix \ref{apx:grammar}. One version implements a prefix and the other one an infix notation. The parser currently supports the prefix notation and correct operator precedences must be handled explicitly with parentheses. To avoid context switches (for developers on Tezos), the syntax is inspired by OCaml \cite{ocaml_docs} and uses the same type notations as Michelson.

\subsection{Code structure}
Assertions are passed to the toolchain as simple text files. This thesis follows the convention of using the file extension \texttt{.tza} for such files (henceforth referred to as ``assertion contracts'') in reference to the file extension \texttt{.tz}used for Michelson source files. Assertion contracts contain at least one assertion for a specific entrypoint of the actual contract (henceforth referred to as ``parent contract''). For the remaining entrypoints the compiler automatically generates empty assertions\footnote{cf. \secref{sec:compiler}}.

An assertion begins with the keyword \texttt{entrypoint} and a signature consisting of an optional tag denoted by \%, followed by a parameter pattern. Parameter patterns are similar to patterns in OCaml --- they are templates that match shapes of data structures and bind identifiers to the components of them. The given parameter pattern must match exactly the parameter type of the entrypoint it is associated with. The body of an assertion can contain a nesting of quantifiers and conditions around exactly one \texttt{assert} expression. Conditions can be used to restrict the quantification domains or for any other constraints. For completeness, the existential quantifier is already included in the grammar, however the current version of the pipeline will reject any assertions containing it.

How the assertions are associated with the respective entrypoint of the parent contract depends on the Smart Contract language. In the case of Solidity, a tag that mirrors the function name is enough to link assertion and entrypoint. In Michelson contracts, on the other hand, entrypoints can be anonymous. Therefore, the assertions are linked to the entrypoints through the tags, the parameter type or both. \secref{sec:typecheck} describes in more detail how the linking is implemented in the Tezos backend.

\subsection{Parameter patterns}
\lstref{lst:patterns} gives an overview of the parameter patterns supported in the assertion language.
\lstinputlisting[caption=Patterns in the assertion language, language=, numbers=none, label=lst:patterns]{listings/patterns.txt}

By using patterns, developers can avoid the explicit handling of special cases, such as empty lists, in the assertion specification. Furthermore, they can be used to put emphasis on the relevant components of the input parameters by replacing the irrelevant ones with wildcards. As an example, the assertion given in \lstref{lst:patterns_usage} concerns the left element of a pair and only applies to non-empty lists. 
\begin{lstlisting}[label=lst:patterns_usage, language=Assertion, numbers=none]
(entrypoint %g (pair (cons a as) _)  ...)
\end{lstlisting}

\subsection{Types and operations}
In order to make the frontend generic, the sets of data types and operations supported by the grammar constitute unions of those present in the target languages. This not only the implementation of a single, common parser, but also allows the transformation to be oblivious of the target platform. As a consequence, the backends are responsible for performing a semantic check and  rejecting any assertions containing types or operations which are not supported by the target language.

The current version of the language contains all data types present in Tezos' VM and a subset of operations that Michelson provides (as of protocol version 007 Delphi). Besides the primitive types \texttt{int}, \texttt{nat} (natural numbers), \texttt{string}, \texttt{bool}, \texttt{bytes} and \texttt{unit}, this includes composite types like \texttt{list}, \texttt{pair}, \texttt{map} or the union type \texttt{or}. Domain specific data types, like \texttt{timestamp} or \texttt{address}, are supported as well. The data type \texttt{fmu} represents a generic fractional monetary unit, which is translated to the corresponding token type of the target platform\footnote{Mutez in Tezos, or Wei in Ethereum}.

\subsection{Example}
\lstinputlisting[caption=Assertion contract for checking if a list is sorted, language=Assertion, label={lst:sorted_assertion}]{listings/sorted.tza}

@@

\section{Accuracy of the approach}\label{sec:accuracy}
\todo{Idea: own chapter "probabilistic model" oder so - OR: model description}
Since the proposed approach implements a probabilistic test, the accuracy as well as the costs incurred by guaranteeing a specified certainty threshold, need to be examined closely. The goal is to identify a formula that, given the domain of a formula in predicate logic, returns an estimate of a lower bound of samples necessary to find an existing counterexample with a probability $p$. From this analysis, it can be derived how effective a formula can be checked on a blockchain with $m$ validators. Sections \ref{sec:coupon} and \ref{sec:prob_threshold} depict the findings from \cite{bernhardt_veigel_2020} and are based on the assumptions that the random generator picks elements independently from a uniform distribution and that there exists exactly one counterexample for the passed parameter.

### 2021-04-17 assertions/doc/thesis-bernhardt/chapters/4-offline_design_tezos.tex
Revision chpt 3 assertion language

\subsection{Type check}
The task of the type checker is to map the assertions from the given file to the respective entrypoints of the parent contract. To this end, the parameter types have to be compared and matched against each other in order to find a correct assignment. The type checker either reads the code of the parent contract from a given file or retrieves the code from a given address on the blockchain. It must be ensured that the created mapping is injective and unambiguous, i.e., each assertion matches exactly one entrypoint and no entrypoint is covered by more than one assertion.\\
As explained in \secref{sec:tezos}, entrypoints can be selected by calling the default entrypoint and wrapping the parameter into the respective union constructors. Alternatively, they can be called explicitly using their tag and passing the parameter as raw value. The same applies for the assertions - by omitting the tag, the parameter type must be declared in respect to the default entrypoint type. If the assertion states an explicit tag, it may declare the raw parameter type. The tags do not necessarily have to be identical with the tags of the parent contract; in some cases, they may be chosen differently for documentation or for readability purposes. However, the tags can and should be used to resolve ambiguity in the mapping between assertions and entrypoints, that is if several entrypoints share the same input type. Ambiguity can also be caused by the fact that entrypoints can be sub-entrypoints of others, thus assertions overlap if a separate assertion is declared for both super- and sub-entrypoint. This is resolved by detecting overlapping assertions during the type check and rejecting assertion contracts where appropriate. \\
As an exemplification, consider a Michelson contract with five entrypoints (including the \texttt{\%default}):
\begin{lstlisting}[numbers=none, language=Michelson]
parameter (or (int %A) (or %BC (int %B) (int %C)))
\end{lstlisting}
The following assertion contract for the given PC contains some valid and some invalid assertions:
\begin{lstlisting}[language=Assertion]
(assertion %A (i : int) ...)             (* valid *)
(assertion %D (i : int) ...)             (* invalid *)
(assertion (right (left (i : int)) ...)  (* valid *)
(assertion %BC (x : (or int int)) ...)   (* invalid *)
\end{lstlisting}
Due to its tag, the fist assertion can be assigned unambiguously to entrypoint A, whereas for the second it is not apparent whether it should be assigned to entrypoint B or C. Since the tag is omitted in the third assertion, the parameter type is given in respect to the default entrypoint and is assigned to B. The last assertion is invalid because it is overlapping with the previous assertion, which was already assigned to entrypoint B.

\subsection{Compiler}
The transformed and type checked assertion AST needs to be retargeted to Michelson code according to the orchestration scheme. Since the modular approach is less complex to implement and avoids having to decompile the parent code, this section describes a compiler that assembles the assertion and contract code according to \figref{fig:modular_assembly}. The target language of the compiler of the pipeline can either be Michelson directly, or another high-level language which supports Michelson as a target. \secref{sec:tezos} lists some examples of such languages. The chosen language then serves as an intermediate representation (IR) of the assertion and manager code, which is ultimately compiled to Michelson by the compiler of the language. After specifying the steps of the compilation in a general way, the compilation approaches are described and discussed in more detail in \secref{sec:direct} and \ref{sec:IR}.

\subsubsection{Compilation of the assertion contracts}
Since the assertions are independent of each other and the manager code, each of them can be compiled in isolation to a separate Michelson program. The generated code of the assertion body and the parameter type can be plugged into a template that already contains the necessary boilerplate code for a contract.

>>>>

\subsection{Type check}\label{sec:typecheck}
The task of the type checker is to map the assertions from the given file to the respective entrypoints of the parent contract. To this end, the parameter types have to be compared and matched against each other in order to find a correct assignment. The type checker either reads the code of the parent contract from a given file or retrieves the code from a given address on the blockchain. It must be ensured that the created mapping is injective and unambiguous, i.e., each assertion matches exactly one entrypoint and no entrypoint is covered by more than one assertion.\\
As explained in \secref{sec:tezos}, entrypoints can be selected by calling the default entrypoint and wrapping the parameter into the respective union constructors. Alternatively, they can be called explicitly using their tag and passing the parameter as raw value. The same applies for the assertions - by omitting the tag, the parameter type must be declared in respect to the default entrypoint type. If the assertion states an explicit tag, it may declare the raw parameter type. The tags do not necessarily have to be identical with the tags of the parent contract; in some cases, they may be chosen differently for documentation or for readability purposes. However, the tags can and should be used to resolve ambiguity in the mapping between assertions and entrypoints, that is if several entrypoints share the same input type. Ambiguity can also be caused by the fact that entrypoints can be sub-entrypoints of others, thus assertions overlap if a separate assertion is declared for both super- and sub-entrypoint. This is resolved by detecting overlapping assertions during the type check and rejecting assertion contracts where appropriate. \\
As an example, consider a Michelson contract with five entrypoints (including the \texttt{\%default}):
\begin{lstlisting}[numbers=none, language=Michelson]
parameter (or (int %A) (or %BC (int %B) (int %C)))
\end{lstlisting}
The following assertion contract for the given PC contains some valid and some invalid assertions:
\begin{lstlisting}[language=Assertion]
(assertion %A (i : int) ...)             (* valid *)
(assertion %D (i : int) ...)             (* invalid *)
(assertion (right (left (i : int)) ...)  (* valid *)
(assertion %BC (x : (or int int)) ...)   (* invalid *)
\end{lstlisting}
Due to its tag, the fist assertion can be assigned unambiguously to entrypoint A, whereas for the second it is not apparent whether it should be assigned to entrypoint B or C. Since the tag is omitted in the third assertion, the parameter type is given in respect to the default entrypoint and is assigned to B. The last assertion is invalid because it is overlapping with the previous assertion, which was already assigned to entrypoint B.

\subsection{Compiler}\label{sec:compiler}
The transformed and type checked assertion AST needs to be retargeted to Michelson code according to the orchestration scheme. Since the modular approach is less complex to implement and avoids having to decompile the parent code, this section describes a compiler that assembles the assertion and contract code according to \figref{fig:modular_assembly}. The target language of the compiler of the pipeline can either be Michelson directly, or another high-level language which supports Michelson as a target. \secref{sec:tezos} lists some examples of such languages. The chosen language then serves as an intermediate representation (IR) of the assertion and manager code, which is ultimately compiled to Michelson by the compiler of the language. After specifying the steps of the compilation in a general way, the compilation approaches are described and discussed in more detail in \secref{sec:direct} and \ref{sec:IR}.

\subsubsection{Compilation of the assertion contracts}
Since the assertions are independent of each other and the manager code, each of them can be compiled in isolation to a separate Michelson program. The generated code of the assertion body and the parameter type can be plugged into a template that already contains the necessary boilerplate code for a contract.
% type inference from patterns
% handling of patterns like cons
